---
title: "CryptoClinical"
author: "Anika Larson"
date: "6/8/2016"
output: pdf_document
---
For this risk estimate, we will use 
[mcmodel](http://www.inside-r.org/packages/cran/mc2d/docs/mcmodel) and 
[evalmcmod](http://www.inside-r.org/packages/cran/mc2d/docs/evalmcmod) from 
the [mc2d](https://cran.r-project.org/web/packages/mc2d/index.html) package. 


```{r}
# Load packages.
library(mc2d)
library(data.table)
```

### Define variables
Set the number of simulations for the variability and uncertainty dimensions. 

```{r}
nsv <- 100
nsu <- 100
ndvar(nsv)  # Variability
ndunc(nsu)   # Uncertainty
rdv <- (nsu/3)
cwv <- (nsu/4)
div.nsv <- (nsv/3)
div5.nsv <- (nsv/5)
div5.nsu <- (nsu/5)
div7.nsu <- (nsu/7)
ns <- nsu * nsv
```

The _mc2d_ functions will set `nsv` to `ndvar()` and `nsu` to `ndunc()` by 
default. Alternatively, we could supply these values to the _mc2d_ model
functions when we call them.

Define the variables we will use to set the `seed` for random sampling and the 
number of `digits` for `print()` statements.

```{r}
seed <- 1
digits <- 5
```

Introducing the bootstrapped dose-response data, fread command and sampling from Brian:

```{r}
alpha <- fread("output_dupont_crypto/alphalist.csv", sep = ',')
n50 <- fread("output_dupont_crypto/N50list.csv", sep = ',')
cdr <- merge(alpha,n50, by = "V1")
scdr <- cdr[sample(nrow(cdr), nsu, replace = TRUE), list(alpha, N50)] #sampling rows in the dose-response bootstrapped table equal to the number of simulations in the uncertainty dimension, aka NSU
```

Leaving that ^ for convenience, but since what we really want is a k-value list made of samples from the three different dose-response estimates (Dupont, Tamu, and UCP):

```{r}
k_dup <- fread("output_crypto_dupont_clinical/klist.csv", sep = ',')
k_tamu <- fread("output_crypto_tamu_clinical/klist.csv", sep = ',')
k_ucp <- fread("output_UCP_crypto_clinical/klist.csv", sep = ',')
kvalues <- rbind(k_dup,k_tamu,k_ucp)
num.kvals <- as.numeric(kvalues$k)
```

Sampling from the Cryptosporidium custom distribution estimates:

```{r}
#Combining the estimates prior to sampling
cw.vl.ln <- fread("concdata_percis.csv", sep = ',')
cw.vlg3 <- unlist(c(cw.vl.ln$c_v1_ln, cw.vl.ln$c_v11_ln, cw.vl.ln$c_v2_ln, cw.vl.ln$c_v5_ln, cw.vl.ln$c_v8a_ln, cw.vl.ln$c_v9a_ln, cw.vl.ln$c_v10_ln))

#Determine which estimate to use: 
cw.vl.samp <- cw.vlg3

```

Sampling from the distribution of the estimates of Crypto reduction by porous ceramic filtration, and converting them from log reduction to percent reduction:
```{r}
  #pfc_reduc set using the distribution
  gcfilt <- fread("cer_filts/CerFiltCrypto.csv", sep = ',')
  colnames(gcfilt) <- c("trial","logr")
  gcfilt$reduc <- 10^(-gcfilt$logr)
  pfc_reduc2 <- as.numeric(gcfilt$reduc)
```

Introducing the daily hosing and showering exposure estimates distribution (discrete normal distributions created in Crystal Ball for # of showers or hosing events per day), and multiplying them by estimated ingestion per event:
```{r}
  hos_show_dists <- fread("exp_params/hosing_showering_discrete_normal_dists.csv", sep = ',')
  #Creating a normal distribution of minutes / shower:
  min_show <- rnorm(mean = 14.644, sd = 11.227, n = 10000)
  #Creating vector of showering exposure values (.001L/min exposure value) :
  shower_vals <- (min_show * hos_show_dists$num_showers * .001)
  
  #Creating a normal distribution of ingestion / hosing event:
  min_hos.ml <- rnorm(mean = 5.500, sd = 6.364, n = 10000)
  min_hos <- min_hos.ml * .001
  #Creating vector of hosing exposure values:
  hosing_vals <- (min_hos * hos_show_dists$hosing_event)
```

### Define exposure model

Within the [mcmodel](http://www.inside-r.org/packages/cran/mc2d/docs/mcmodel)
function, use [mcstoc](http://www.inside-r.org/packages/cran/mc2d/docs/mcstoc)
to define "mc nodes" for each component of the model. 

For each stochastic variable, use the `mcstoc()` function to create a 
"mcnode", supply a probability function, the node type as "V" for variability 
or "U" for uncertainty, and any additional parameters to be passed to the 
probability function. 

For deterministic variables, create nodes with the `mcdata()` function 
using the "0" type.

We will model the deterministic factors as uniform probablity distributions.

The last statement makes an "mc" object using the 
[mc](http://www.inside-r.org/packages/cran/mc2d/docs/mc) function.

```{r}
# Define an exposure model for evaluation by evalmcmod().
expo.mod1 <- mcmodel({
  # Values from Example 6.18 from Quantitative Microbial Risk Assessment, 
  # 2nd Edition by Charles N. Haas, Joan B. Rose, and Charles P. Gerba. 
  # (Wiley, 2014), pp. 215-216. Other fictitious values are noted below.
  
  #Cistern water viral loading (viruses/L):
  cw.vl2 <- mcstoc(rempiricalD, type = "VU", values = cw.vl.samp)
  cw.vl <- cw.vl2
  
  # Water consumption (L/day):
  dw.ing1 <- mcdata(1.466, type = "U")
  dw.ing2 <- mcstoc(rpois, type = "V", lambda = 1.466)
  dw.ing <- dw.ing1
  
  #Daily hosing exposure sampling from distribution
  daily_hose_exp <- mcstoc(rempiricalD, type = "V", values = hosing_vals, rtrunc = TRUE, linf = 0)
  hosing <- daily_hose_exp

  #Daily showering sampling from distribution:
  daily_shower_exp <- mcstoc(rempiricalD, type = "V", values = shower_vals, rtrunc = TRUE, linf = 0)
  shower <- daily_shower_exp
  
  # Estimate the exposure based on drinking water, hosing and ingestion, assuming no particular distribution:
  exporaw.mc1 <- (cw.vl * (dw.ing + hosing + shower))
  
    #Reduction by one of the three estimates in Table 3, assuming all are equally likely:
    exporaw.freechl.tog <- mcstoc(rempiricalD, values = c(.53,.5096,.01), prob = c(.33,33,33), type = "V")
    exporaw_free_chl <- (exporaw.mc1 * exporaw.freechl.tog)
  
  #Reduction by porous ceramic filtration based on the distribution entered earlier:
  pfc_reduc <- mcstoc(rempiricalD, type = "V", values = pfc_reduc2)
  exporaw.porous_ceram <- (exporaw.mc1 * pfc_reduc)
  
  #Reduction by porous ceramic filtration, assuming 99% reduction:
  exporaw.por_ceram_singval <- (exporaw.mc1 * .01)
      
  #Reduction by UV filtration:
  exporaw.uv <- (exporaw.mc1 * .0001)

  #Final exposure - can be assigned to whichever filtration system is in use:
  expo.mc1 <-  exporaw.porous_ceram
  
  dose1 <- mcstoc(rpois, type = "VU", lambda = expo.mc1)
  dose2 <- expo.mc1
  dose <- dose2
  
  # Create mcnodes from previously sampled dose-response bootstrap estimates - a and n50 just left here for convenience.
  a <- mcdata(scdr$alpha, type = "U")
  n50 <- mcdata(scdr$N50, type = "U")
  k1 <- mcstoc(rempiricalD, type = "V", values = num.kvals)
  k <- k1
  
  #estimating risk with exponential equation:
  risk1 <- (1 - exp(-dose*k))
  risk2 <- mcstoc(rpois, type = "VU", lambda = risk1)
  risk <- risk1

  #yearlyrisk
  yearrisk <- 1- ((1 - risk)^365)
  
    #RAW VERSION:
  #Either directly consider the exposure the dose, or add uncertainty by considering it a Poisson distribution centered at the previous estimate (? - this showed up in an earlier version) :
  dose1.raw <- mcstoc(rpois, type = "VU", lambda = exporaw.mc1)
  dose2.raw <- exporaw.mc1
  dose.raw <- dose2.raw
  
  #this equation is my version of what came off QMRA Wiki <http://qmrawiki.canr.msu.edu/index.php?title=Dose_response_assessment> and someone should check it.  BUT it's beta-poisson, and just left here for convenience.
  riskbp <- 1 - (1 + dose*((2^(1/a)-1)/n50))^(-a) 
  
      #Exponential version of the equation (also from QMRA Wiki and please lmk if it looks wrong):
  risk1.raw <- (1 - exp(-dose.raw*k))
  risk2 <- mcstoc(rpois, type = "VU", lambda = risk1)
  risk.raw <- risk1.raw
  
  #yearlyrisk
  yearrisk.raw <- 1- ((1 - risk.raw)^365)
  
  #FREE CHLORINE VERSION:
  #Either directly consider the exposure the dose, or add uncertainty by considering it a Poisson distribution centered at the previous estimate (? - this showed up in an earlier version) :
  dose1 <- mcstoc(rpois, type = "VU", lambda = exporaw_free_chl)
  dose2 <- exporaw_free_chl
  dose.chl <- dose2
  
  #Exponential version of the equation (also from QMRA Wiki and please lmk if it looks wrong):
  risk1.chl <- (1 - exp(-dose.chl*k))
  risk2 <- mcstoc(rpois, type = "VU", lambda = risk1)
  risk.chl <- risk1.chl
  
  #yearlyrisk
  yearrisk.chl <- 1- ((1 - risk.chl)^365)
  
    #UV VERSION:
  dose.uv <- exporaw.uv
  #Exponential version of the equation:
  risk1.uv <- (1 - exp(-dose.uv*k))
  risk.uv <- risk1.uv
  
  #yearlyrisk
  yearrisk.uv <- 1- ((1 - risk.uv)^365)
  
      #POROUS CERAMIC FILTRATION VERSION:
  dose2.pcf <- exporaw.porous_ceram
  dose.pcf <- dose2.pcf
  
      #Exponential version of the equation (also from QMRA Wiki and please lmk if it looks wrong):
  risk1.pcf <- (1 - exp(-dose.pcf*k))
  risk2.pcf <- mcstoc(rpois, type = "VU", lambda = risk1.pcf)
  risk.pcf <- risk1.pcf
  
  #yearlyrisk
  yearrisk.pcf <- 1- ((1 - risk.pcf)^365)
  
  # Build a mc model from all of the mcnode objects.
  mc(cw.vl2, cw.vl, dw.ing, daily_hose_exp, daily_shower_exp, exporaw_free_chl, exporaw.uv,  exporaw.porous_ceram, pfc_reduc,  a, n50, k1, k, expo.mc1, dose, risk1, risk, yearrisk, risk.raw, dose.raw, yearrisk.raw,risk.chl, dose.chl, yearrisk.chl,risk.uv, dose.uv, yearrisk.uv,risk.pcf, dose.pcf, yearrisk.pcf) 
})
```

### Evaluate the model

Evaluate the model with 401 iterations in the variability dimension and 59 
iterations in the uncertainty dimension, as set previously.

```{r}
expo.ev1 <- evalmcmod(expo.mod1, seed = seed)
print(expo.ev1, digits = digits)
```

### Summarize results

Print a summary and a plot of the evaluation results (`expo.ev1`). 

```{r results-ev1}
# Summarize the results.
summary(expo.ev1)

# Plot the results.
plot(expo.ev1)
```

Report the median and median of the exposure medians with a 95% confidence interval (CI95). 

```{r}
median.expo <- sapply(1:ndunc(), function(j) median(expo.ev1$expo.mc1[, j, ]))
median(median.expo)
quantile(median.expo, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```


RAW:

Report the median and median of the DAILY risk probability medians with a 95% confidence interval (CI95). 

```{r median median-risk.raw}
median.risk.raw <- sapply(1:ndunc(), function(j) median(expo.ev1$risk.raw[, j, ]))
median(median.risk.raw)
quantile(median.risk.raw, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the median and median of the YEARLY risk probability medians with a 95% confidence interval (CI95). 

```{r median risk.raw}
median.yearrisk.raw <- sapply(1:ndunc(), function(j) median(expo.ev1$yearrisk.raw[, j, ]))
median(median.yearrisk.raw)
quantile(median.yearrisk.raw, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Plot the empirical cumulative distribution function (ecdf) of the risk model 
(`risk`) estimates.

```{r plot-risk.raw}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.raw)
```

Label the plot:

```{r plot-riskwlab.raw}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.raw, main = "", ylab = "Proportion of Estimates", xlab = "Daily Probability of Infection")
title("Daily Infec. Risk with Cryptosporidium by RCRW
from Raw Water", cex.main = 1, font.main= 1, line = 7)
```


```{r plot-yearriskwlab.raw}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$yearrisk.raw, main = "", ylab = "Proportion of Estimates", xlab = "Yearly Probability of Infection")
title("Yearly Infec. Risk with Cryptosporidium by RCRW
from Raw Water", cex.main = 1, font.main= 1, line = 7)
```

CHLORINATED:

Report the median and median of the DAILY risk probability medians with a 95% confidence interval (CI95). 

```{r median median-risk.chl}
median.risk.chl <- sapply(1:ndunc(), function(j) median(expo.ev1$risk.chl[, j, ]))
median(median.risk.chl)
quantile(median.risk.chl, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the median and median of the YEARLY risk probability medians with a 95% confidence interval (CI95). 

```{r median risk.chl}
median.yearrisk.chl <- sapply(1:ndunc(), function(j) median(expo.ev1$yearrisk.chl[, j, ]))
median(median.yearrisk.chl)
quantile(median.yearrisk.chl, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Plot the empirical cumulative distribution function (ecdf) of the risk model 
(`risk`) estimates.

```{r plot-risk.chl}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.chl)
```

Label the plot:

```{r plot-riskwlab.chl}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.chl, main = "", ylab = "Proportion of Estimates", xlab = "Daily Probability of Infection")
title("Daily Infec. Risk with Cryptosporidium by RCRW
from Chlorinated Water", cex.main = 1, font.main= 1, line = 7)
```


```{r plot-yearriskwlab.chl}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$yearrisk.chl, main = "", ylab = "Proportion of Estimates", xlab = "Yearly Probability of Infection")
title("Yearly Infec. Risk with Cryptosporidium by RCRW
from Chlorinated Water", cex.main = 1, font.main= 1, line = 7)
```

UV TREATED:

Report the median and median of the DAILY risk probability medians with a 95% confidence interval (CI95). 

```{r median median-risk.uv}
median.risk.uv <- sapply(1:ndunc(), function(j) median(expo.ev1$risk.uv[, j, ]))
median(median.risk.uv)
quantile(median.risk.uv, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the median and median of the YEARLY risk probability medians with a 95% confidence interval (CI95). 

```{r median risk.uv}
median.yearrisk.uv <- sapply(1:ndunc(), function(j) median(expo.ev1$yearrisk.uv[, j, ]))
median(median.yearrisk.uv)
quantile(median.yearrisk.uv, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Plot the empirical cumulative distribution function (ecdf) of the risk model 
(`risk`) estimates.

```{r plot-risk.uv}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.uv)
```

Label the plot:

```{r plot-riskwlab.uv}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.uv, main = "", ylab = "Proportion of Estimates", xlab = "Daily Probability of Infection")
title("Daily Infec. Risk with Cryptosporidium by RCRW
from UV Treated Water", cex.main = 1, font.main= 1, line = 7)
```


```{r plot-yearriskwlab.uv}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$yearrisk.uv, main = "", ylab = "Proportion of Estimates", xlab = "Yearly Probability of Infection")
title("Yearly Infec. Risk with Cryptosporidium by RCRW
from UV Treated Water", cex.main = 1, font.main= 1, line = 7)
```

POROUS CERAMIC FILTRATION:

Report the median and median of the DAILY risk probability medians with a 95% confidence interval (CI95). 

```{r median median-risk.pcf}
median.risk.pcf <- sapply(1:ndunc(), function(j) median(expo.ev1$risk.pcf[, j, ]))
median(median.risk.pcf)
quantile(median.risk.pcf, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the median and median of the YEARLY risk probability medians with a 95% confidence interval (CI95). 

```{r median risk.pcf}
median.yearrisk.pcf <- sapply(1:ndunc(), function(j) median(expo.ev1$yearrisk.pcf[, j, ]))
median(median.yearrisk.pcf)
quantile(median.yearrisk.pcf, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Plot the empirical cumulative distribution function (ecdf) of the risk model 
(`risk`) estimates.

```{r plot-risk.pcf}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.pcf)
```

Label the plot:

```{r plot-riskwlab.pcf}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk.pcf, main = "", ylab = "Proportion of Estimates", xlab = "Daily Probability of Infection")
title("Daily Infec. Risk with Cryptosporidium by RCRW
from Water Treated with PCF", cex.main = 1, font.main= 1, line = 7)
```


```{r plot-yearriskwlab.pcf}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$yearrisk.pcf, main = "", ylab = "Proportion of Estimates", xlab = "Yearly Probability of Infection")
title("Yearly Infec. Risk with Cryptosporidium by RCRW
from Water Treated with PCF", cex.main = 1, font.main= 1, line = 7)
```



Manually remake the graph, this time using yearrisk, which should just be another mcnode object with all values between 0 and 1:

The first one plots yearrisk without a log x scale:
```{r plot-withyearrisk}
library(reshape)
library(ggplot2)
Risk.mat.1 <- expo.ev1$yearrisk[,,]
probs1 <- c(0.025, 0.25, 0.50, 0.75, 0.975)
quant1 <- as.data.frame(t(apply(Risk.mat.1, 1, quantile, probs = probs1)))
ecdfs1 <- sapply(names(quant1), function(q) ecdf(quant1[[q]]))
plot(ecdfs1[['50%']], main = '')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
lines <- mapply(function(e, g) lines(e, col = g), ecdfs1, grays)
```


Plotting yearrisk ECDF plot with ggplot2:

```{r plot-withyearriskgpplot}
quant.melt1 <- suppressMessages(melt(quant1))
names(quant.melt1) <- c('q', 'x')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
ggplot(quant.melt1, aes(x = x)) + theme_bw() + theme(legend.position = 'none') + geom_hline(yintercept = c(0, 1), linetype = "dashed", color = 'gray') + stat_ecdf(aes(group = q, color = q)) + xlab('x') + ylab('Fn(x)') + scale_colour_manual(values = grays)
```

Plot Version 2 with ggplot2, and a log-scale x axis:

```{r plot-withyearrisklogscale}
quant.melt1 <- suppressMessages(melt(quant1))
names(quant.melt1) <- c('q', 'x')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
ggplot(quant.melt1, aes(x = x)) + theme_bw() + theme(legend.position = 'none') + geom_hline(yintercept = c(0, 1), linetype = "dashed", color = 'gray') + stat_ecdf(aes(group = q, color = q)) + xlab('x') + ylab('Fn(x)') + scale_colour_manual(values = grays) + scale_x_log10()
```
