---
title: "Giardia_riskdraft"
author: "Anika Larson"
date: "5/19/2016"
output: html_document
---

To Do:
Rewrite this for an exponential equation for Giardia, using new Giardia dose-response relationship data.

For this risk estimate, we will use 
[mcmodel](http://www.inside-r.org/packages/cran/mc2d/docs/mcmodel) and 
[evalmcmod](http://www.inside-r.org/packages/cran/mc2d/docs/evalmcmod) from 
the [mc2d](https://cran.r-project.org/web/packages/mc2d/index.html) package. 


```{r warning=FALSE}
# Load packages.
library(mc2d)
library(data.table)
library(sfsmisc)
library(graphics)
```

### Define variables

Set the number of simulations for the variability and uncertainty dimensions. 

```{r message=FALSE}
nsv <- 401
nsu <- 108
ndvar(nsv)  # Variability
ndunc(nsu)   # Uncertainty
```

The _mc2d_ functions will set `nsv` to `ndvar()` and `nsu` to `ndunc()` by 
default. Alternatively, we could supply these values to the _mc2d_ model
functions when we call them.

Define the variables we will use to set the `seed` for random sampling and the 
number of `digits` for `print()` statements.

```{r}
seed <- 1
digits <- 5
```

Introducing the bootstrapped dose-response data, fread command and sampling from Brian:

```{r}
klist <- fread("output_rendtorff_giardia/klist.csv", sep = ',')
gdrk <- klist[sample(nrow(klist), nsu, replace = TRUE), list(x)] #sampling rows in the dose-response bootstrapped table equal to the number of simulations in the uncertainty dimension, aka NSU
```

Sampling from the Giardia Crabtree custom distribution estimates:

```{r}
cw.vlg <- fread("cwvl_estimates/giardia_estwol.csv", sep = ',')
cw.vl.samp <- sample(cw.vlg[,c(V1)], nsu, replace = TRUE) 
```

```{r}
  #pfc_reduc set using the distribution
  gcfilt <- fread("cer_filts/CerFiltGiardia.csv", sep = ',')
  colnames(gcfilt) <- c("trial","logr")
  gcfilt$reduc <- (1 - (10^(-gcfilt$logr)))
  pfc_reduc2 <- gcfilt[sample(nrow(gcfilt), nsu, replace = TRUE), list(reduc)]
```

  
We will use this variable to explicitly set the seed with the various
_mc2d_ functions. Another approach would be to only set it through the
model evaluation functions, or not at all. Since we want to do our best
to provide the most reproducible results, we set the seed explicitly.

### Define exposure model

Within the [mcmodel](http://www.inside-r.org/packages/cran/mc2d/docs/mcmodel)
function, use [mcstoc](http://www.inside-r.org/packages/cran/mc2d/docs/mcstoc)
to define "mc nodes" for each component of the model. 

For each stochastic variable, use the `mcstoc()` function to create a 
"mcnode", supply a probability function, the node type as "V" for variability 
or "U" for uncertainty, and any additional parameters to be passed to the 
probability function. 

For deterministic variables, create nodes with the `mcdata()` function 
using the "0" type.

We will model the deterministic factors as uniform probablity distributions.

The last statement makes an "mc" object using the 
[mc](http://www.inside-r.org/packages/cran/mc2d/docs/mc) function.

```{r}
# Define an exposure model for evaluation by evalmcmod().
expo.mod1 <- mcmodel({
  # Values from Example 6.18 from Quantitative Microbial Risk Assessment, 
  # 2nd Edition by Charles N. Haas, Joan B. Rose, and Charles P. Gerba. 
  # (Wiley, 2014), pp. 215-216. Other fictitious values are noted below.
  
   # Cistern water viral loading (viruses/L):
  cw.vl <- mcdata(cw.vl.samp, type = "U")
  
  # Water consumption (L/day):
  dw.ing <- mcdata(1.466, type = "0")

  # Hosing per event ingestion rate (L/event): worst case estimate = 10mL or .01L per day
  hose.perevent <- mcdata(.01, type = "0")

  # Hosing frequency (hosing/day), fictitious discrete distribution (this basically says that there is a 10% chance someone never hoses, a 70% chance they use the hose every other day, and a 10% chance they use the hose every day) - will be IGNORED for the moment:
  hosing.frequency <- mcstoc(rempiricalD, type = "V", seed = seed, 
                             values = c(0, 0.5, 1), 
                             prob = c(0.1, 0.7, 0.2))
  
   # Showering per event ingestion rate (L/event): worst case estimate = 10mL or .01L per day
  shower.perevent <- mcdata(.01, type = "0")
  
  # Estimate the exposure based ONLY on drinking water ingestion using the 0, V and U nodes to create a VU node.
  exporaw.mc1 <- (cw.vl * (dw.ing + hose.perevent + shower.perevent)) 
  
  #Reduction by free chlorine at 5ppm - this is based on the "range" of values from 99% reduction to 99.8% reduction, but can be changed to just a single value:
  fc_reduc <- mcstoc(rempiricalD, type = "V", seed = seed, values = c(0.01, 0.007, 0.004, 0.002), prob = c(0.25, 0.25, 0.25, 0.25))
  exporaw.5ppm_fcl <- (exporaw.mc1 * fc_reduc)
  
  #Reduction by UV: 
  exporaw.uv <- (exporaw.mc1 * .0001)
  
    #Reduction by porous ceramic filtration - this is based on the "range" of values from 98.7% reduction to 99.8% reduction, but can be changed to just a single value:
  pfc_reduc1 <- mcstoc(rempiricalD, type = "V", seed = seed, values = c(0.013, 0.009, 0.006, 0.002), prob = c(0.25, 0.25, 0.25, 0.25))

  pfc_reduc <- mcdata(pfc_reduc2$reduc, type = "U")
  exporaw.pfc_reduc <- (exporaw.mc1 * pfc_reduc)
  
  #Reduction by porous ceramic filtration AND free chlorine 5ppm:
  exporaw.fcl_pfc <- (exporaw.mc1 * pfc_reduc * fc_reduc)
  
    #Final exposure - just make this whichever one you want to use at the moment:
  expo.mc1 <- exporaw.pfc_reduc
  
  #Estimate the risk using a dose-response model pulling from a bootstrapped model of the dose-response
  dose <- mcstoc(rpois, type = "VU", lambda = expo.mc1)

  k <- mcdata(gdrk$x, type = "U")
  
  risk <- (1 - exp(-dose*k)) #this equation is my version an exponential DR equation from what came off QMRA Wiki <http://qmrawiki.canr.msu.edu/index.php?title=Dose_response_assessment> and someone should check it -A
  
  #yearlyrisk
  yearrisk <- 1 - ((1 - risk)^365)
  
  logyearrisk <- log10(yearrisk)
  
  # Build a mc model from all of the mcnode objects.
  mc(cw.vl, dw.ing, hose.perevent, shower.perevent, k, exporaw.mc1, exporaw.5ppm_fcl, fc_reduc, pfc_reduc, exporaw.pfc_reduc, exporaw.uv, expo.mc1, risk, yearrisk, logyearrisk) 
})
```

### Evaluate the model

Evaluate the model with 5000 iterations in the variability dimension and 250 
iterations in the uncertainty dimension, as set previously.

```{r}
expo.ev1 <- evalmcmod(expo.mod1, seed = seed)
print(expo.ev1, digits = digits)
```

### Summarize results

Print a summary and a plot of the evaluation results (`expo.ev1`). 

```{r results-ev1}
# Summarize the results.
summary(expo.ev1)

# Plot the results.
plot(expo.ev1)
```


Report the mean and median of the DAILY risk probability means with a 95% confidence interval (CI95). 

```{r mean median-risk}
mean.risk <- sapply(1:ndunc(), function(j) mean(expo.ev1$risk[, j, ]))
mean(mean.risk)
quantile(mean.risk, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the mean and median of the YEARLY risk probability means with a 95% confidence interval (CI95). 

```{r mean risk}
mean.yearrisk <- sapply(1:ndunc(), function(j) mean(expo.ev1$yearrisk[, j, ]))
mean(mean.yearrisk)
quantile(mean.yearrisk, probs = seq(0, 1, 0.025))[c("2.5%", "25%", "50%", "75%", "97.5%")]
```


Plot the empirical cumulative distribution function (ecdf) of the exposure model 
(`expo.mc1`) estimates.

```{r plot-mc1}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$expo.mc1)
```

Plot the empirical cumulative distribution function (ecdf) of the risk model 
(`risk`) estimates.

```{r plot-risk}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk)
```

So if we did this right, could the plot above can be correctly labeled as follows?



```{r plot-riskwlab}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk, main = "", ylab = "Proportion of Estimates", xlab = "Daily Probability of Infection")
title("Daily Infec. Risk with Giardia by RCRW
in the USVI", cex.main = 1, font.main= 1, line = 7)
```


```{r plot-yearriskwlab}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$yearrisk, main = "", ylab = "Proportion of Estimates", xlab = "Yearly Probability of Infection")
title("Yearly Infec. Risk with Giardia by RCRW
in the USVI", cex.main = 1, font.main= 1, line = 7)
```



Manually remake the graph, this time using yearrisk, which should just be another mcnode object with all values between 0 and 1:

The first one plots yearrisk without a log x scale:
```{r plot-withyearrisk}
library(reshape)
library(ggplot2)
Risk.mat.1 <- expo.ev1$yearrisk[,,]
probs1 <- c(0.025, 0.25, 0.50, 0.75, 0.975)
quant1 <- as.data.frame(t(apply(Risk.mat.1, 1, quantile, probs = probs1)))
ecdfs1 <- sapply(names(quant1), function(q) ecdf(quant1[[q]]))
plot(ecdfs1[['50%']], main = '')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
lines <- mapply(function(e, g) lines(e, col = g), ecdfs1, grays)
```


Plotting yearrisk ECDF plot with ggplot2:

```{r plot-withyearriskgpplot}
quant.melt1 <- suppressMessages(melt(quant1))
names(quant.melt1) <- c('q', 'x')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
ggplot(quant.melt1, aes(x = x)) + theme_bw() + theme(legend.position = 'none') + geom_hline(yintercept = c(0, 1), linetype = "dashed", color = 'gray') + stat_ecdf(aes(group = q, color = q)) + xlab('x') + ylab('Fn(x)') + scale_colour_manual(values = grays)
```

Plot Version 2 with ggplot2, and a log-scale x axis:

```{r plot-withyearrisklogscale}
quant.melt1 <- suppressMessages(melt(quant1))
names(quant.melt1) <- c('q', 'x')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
ggplot(quant.melt1, aes(x = x)) + theme_bw() + theme(legend.position = 'none') + geom_hline(yintercept = c(0, 1), linetype = "dashed", color = 'gray') + stat_ecdf(aes(group = q, color = q)) + xlab('x') + ylab('Fn(x)') + scale_colour_manual(values = grays) + scale_x_log10()
```

