---
title: "cryptosporidium_riskdraft"
author: "Anika Larson"
date: "5/10/2016"
output: html_document
---
For this risk estimate, we will use 
[mcmodel](http://www.inside-r.org/packages/cran/mc2d/docs/mcmodel) and 
[evalmcmod](http://www.inside-r.org/packages/cran/mc2d/docs/evalmcmod) from 
the [mc2d](https://cran.r-project.org/web/packages/mc2d/index.html) package. 


```{r warning=FALSE}
# Load packages.
library(mc2d)
library(data.table)
```

### Define variables

Set the number of simulations for the variability and uncertainty dimensions. 

```{r message=FALSE}
nsv <- 401
nsu <- 108
ndvar(nsv)  # Variability
ndunc(nsu)   # Uncertainty
rdv <- (nsu/3)
cwv <- (nsu/4)
ns <- nsu * nsv
```

The _mc2d_ functions will set `nsv` to `ndvar()` and `nsu` to `ndunc()` by 
default. Alternatively, we could supply these values to the _mc2d_ model
functions when we call them.

Define the variables we will use to set the `seed` for random sampling and the 
number of `digits` for `print()` statements.

```{r}
seed <- 1
digits <- 5
```

Introducing the bootstrapped dose-response data, fread command and sampling from Brian:

```{r}
alpha <- fread("output_dupont_crypto/alphalist.csv", sep = ',')
n50 <- fread("output_dupont_crypto/N50list.csv", sep = ',')
cdr <- merge(alpha,n50, by = "V1")
scdr <- cdr[sample(nrow(cdr), nsu, replace = TRUE), list(alpha, N50)] #sampling rows in the dose-response bootstrapped table equal to the number of simulations in the uncertainty dimension, aka NSU
```

Leaving that ^ for convenience, but since what we really want is a k-value list made of samples from the three different dose-response estimates (Dupont, Tamu, and UCP):

```{r}
k_dup <- fread("output_dupont_crypto/klist_dupont.csv", sep = ',')
gdrkD <- k_dup[sample(nrow(k_dup), rdv, replace = TRUE), list(k)] #sampling rows in the dose-response bootstrapped table equal to 1/3 the number of simulations in the uncertainty dimension, aka NSU
k_tamu <- fread("output_tamu_crypto/klist_tamu.csv", sep = ',')
gdrkT <- k_tamu[sample(nrow(k_tamu), rdv, replace = TRUE), list(k)]
k_ucp <- fread("output_UCP_crypto/klist_ucp.csv", sep = ',')
gdrkU <- k_ucp[sample(nrow(k_ucp), rdv, replace = TRUE), list(k)]
kval <- unlist(c(gdrkD,gdrkT,gdrkU))
```


Sampling from the Cryptosporidium custom distribution estimates:

```{r}
cw.vlg <- fread("cwvl_estimates/crypto_est.csv", sep = ',')
cw.vl.samp <- sample(cw.vlg[,c(V1)], nsu, replace = TRUE) #sampling rows in the dose-response bootstrapped table equal to the number of simulations in the uncertainty dimension, aka NSU
```


```{r}
  #pfc_reduc set using the distribution
  gcfilt <- fread("cer_filts/CerFiltCrypto.csv", sep = ',')
  colnames(gcfilt) <- c("trial","logr")
  gcfilt$reduc <- (1 - (10^(-gcfilt$logr)))
  pfc_reduc2 <- gcfilt[sample(nrow(gcfilt), nsu, replace = TRUE), list(reduc)]
```

```{r}
  #Creating a distribution for hosing
  hose <- fread("exp_params/Hosing.csv", sep = ',')
  colnames(hose) <- c("trial","num", "mL")
  hose$hose_exp1 <- ((hose$num * hose$mL) / 7)
  hose_exp <- hose[sample(nrow(hose), nsu, replace = TRUE), list(hose_exp1)]
```

```{r}
  #Creating a distribution for showering
  shower <- fread("exp_params/Showers.csv", sep = ',')
  colnames(shower) <- c("trial","num", "mL")
  shower$shower_exp1 <- ((shower$num * shower$mL) / 7)
  shower_exp <- shower[sample(nrow(shower), nsu, replace = TRUE), list(shower_exp1)]
```
### Define exposure model

Within the [mcmodel](http://www.inside-r.org/packages/cran/mc2d/docs/mcmodel)
function, use [mcstoc](http://www.inside-r.org/packages/cran/mc2d/docs/mcstoc)
to define "mc nodes" for each component of the model. 

For each stochastic variable, use the `mcstoc()` function to create a 
"mcnode", supply a probability function, the node type as "V" for variability 
or "U" for uncertainty, and any additional parameters to be passed to the 
probability function. 

For deterministic variables, create nodes with the `mcdata()` function 
using the "0" type.

We will model the deterministic factors as uniform probablity distributions.

The last statement makes an "mc" object using the 
[mc](http://www.inside-r.org/packages/cran/mc2d/docs/mc) function.

```{r}
# Define an exposure model for evaluation by evalmcmod().
expo.mod1 <- mcmodel({
  # Values from Example 6.18 from Quantitative Microbial Risk Assessment, 
  # 2nd Edition by Charles N. Haas, Joan B. Rose, and Charles P. Gerba. 
  # (Wiley, 2014), pp. 215-216. Other fictitious values are noted below.
  
   # Cistern water viral loading (viruses/L):
  cw.vl1 <- mcdata(cw.vl.samp, type = "U")
  cw.vl2 <- mcstoc(rempiricalD, type = "V", values = cw.vl.samp)
  cw.vl <- cw.vl2
  
  # Water consumption (L/day):
  dw.ing1 <- mcdata(1.466, type = "U")
  dw.ing2 <- mcstoc(rpois, type = "V", lambda = 1.466)
  dw.ing <- dw.ing1
  
  # Hosing per event ingestion rate (L/event): worst case estimate = 10mL or .01L per day
  hose.perevent <- mcdata(.01, type = "U")
  hose.perevent.dis <- mcdata(hose_exp$hose_exp1, type = "U")
  #Daily hosing exposure sampling from distribution
  daily_hose_exp <- mcstoc(rempiricalD, type = "V", values = hose_exp$hose_exp1)
  hosing <- daily_hose_exp
  
   # Showering per event ingestion rate (L/event): worst case estimate = 10mL or .01L per day
  shower.perevent <- mcdata(.01, type = "U")
  shower.perevent.dis <- mcdata(shower_exp$shower_exp1, type = "U")
  
  #Daily showering sampling from distribution:
  daily_shower_exp <- mcstoc(rempiricalD, type = "V", values = shower_exp$shower_exp1)
  shower <- daily_shower_exp
  
  # Estimate the exposure based on drinking water, hosing and ingestion, assuming no particular distribution:
  exporaw.mc1 <- (cw.vl * (dw.ing + hosing + shower))
  
  #Reduction by 2ppm free chlorine:
    exporaw.2ppm_free_chl1 <- (exporaw.mc1 * .4904)
    exporaw_2ppm_unc <- mcdata(.4904, type = "U")
    exporaw_2ppm_free_chl2 <- (exporaw.mc1 * exporaw_2ppm_unc)
    exporaw.2ppm_free_chl <- exporaw_2ppm_free_chl2
  
  #Reduction by 15ppm free chlorine:
    exporaw.15ppm_free_chl1 <- (exporaw.mc1 * .47)
    exporaw_15ppm_unc <- mcdata(.47, type = "U")
    exporaw_15ppm_free_chl2 <- (exporaw.mc1 * exporaw_15ppm_unc)
    exporaw.15ppm_free_chl <- exporaw_15ppm_free_chl2
    
    #Reduction by one or the other, assuming each is added half the time:
    exporaw.freechl.tog <- mcstoc(rempiricalD, values = c(.47,.4904), prob = c(.5,.5), type = "V")
    exporaw_free_chl <- (exporaw.mc1 * exporaw.freechl.tog)
  
  #Reduction by porous ceramic filtration - this is based on the "range" of values from 66.1% reduction to 99.6% reduction:
  pfc_reduc3 <- mcdata(pfc_reduc2$reduc, type = "U")
  pfc_reduc4 <- mcstoc(rempiricalC, type = "V", values = pfc_reduc2$reduc, min=0, max=1)
  pfc_reduc <- pfc_reduc4
  exporaw.porous_ceram <- (exporaw.mc1 * pfc_reduc)
  
  #Reduction by porous ceramic filtration, assuming 99% reduction:
  exporaw.por_ceram_singval <- (exporaw.mc1 * .01)
      
  #Reduction by UV filtration:
  exporaw.uv <- (exporaw.mc1 * .0001)

  #Final exposure - just make this whichever one you want to use at the moment:
  expo.mc1 <-  exporaw.por_ceram_singval

  
  #Estimate the risk using a dose-response model pulling from a bootstrapped model of the dose-response
  dose1 <- mcstoc(rpois, type = "VU", lambda = expo.mc1)
  dose2 <- expo.mc1
  dose <- dose2
  
  # Create mcnodes from previously sampled dose-response bootstrap estimates. - Brian
  
  a <- mcdata(scdr$alpha, type = "U")
  n50 <- mcdata(scdr$N50, type = "U")
  
  #this equation is my version of what came off QMRA Wiki <http://qmrawiki.canr.msu.edu/index.php?title=Dose_response_assessment> and someone should check it.  BUT it's beta-poisson, and just left here for convenience.
  riskbp <- 1 - (1 + dose*((2^(1/a)-1)/n50))^(-a) 
  
  #Exponential version of the equation (also from QMRA Wiki and please lmk if it looks wrong):
  k1 <- mcdata(kval, type = "U")
  k2 <- mcstoc(rempiricalD, type = "VU", values = kval)
  k <- k2
  
  risk1 <- (1 - exp(-dose*k))
  risk2 <- mcstoc(rpois, type = "VU", lambda = risk1)
  risk <- risk1
  
  
  #yearlyrisk
  yearrisk <- 1- ((1 - risk)^365)
  
  # Build a mc model from all of the mcnode objects.
  mc(cw.vl1, cw.vl2, cw.vl, dw.ing, hose.perevent, shower.perevent, exporaw.15ppm_free_chl, exporaw_free_chl, exporaw.uv,  exporaw.porous_ceram, pfc_reduc,  a, n50, k1, k2, k, expo.mc1, dose, riskbp, risk1, risk, yearrisk) 
})
```

### Evaluate the model

Evaluate the model with 401 iterations in the variability dimension and 59 
iterations in the uncertainty dimension, as set previously.

```{r}
expo.ev1 <- evalmcmod(expo.mod1, seed = seed)
print(expo.ev1, digits = digits)
```

### Summarize results

Print a summary and a plot of the evaluation results (`expo.ev1`). 

```{r results-ev1}
# Summarize the results.
summary(expo.ev1)

# Plot the results.
plot(expo.ev1)
```

Report the mean and median of the exposure means with a 95% confidence interval (CI95). 

```{r}
mean.expo <- sapply(1:ndunc(), function(j) mean(expo.ev1$expo.mc1[, j, ]))
mean(mean.expo)
quantile(mean.expo, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the mean and median of the DAILY risk probability means with a 95% confidence interval (CI95). 

```{r mean median-risk}
mean.risk <- sapply(1:ndunc(), function(j) mean(expo.ev1$risk[, j, ]))
mean(mean.risk)
quantile(mean.risk, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Report the mean and median of the YEARLY risk probability means with a 95% confidence interval (CI95). 

```{r mean risk}
mean.yearrisk <- sapply(1:ndunc(), function(j) mean(expo.ev1$yearrisk[, j, ]))
mean(mean.yearrisk)
quantile(mean.yearrisk, probs = seq(0, 1, 0.025))[c("2.5%", "50%", "97.5%")]
```

Plot the empirical cumulative distribution function (ecdf) of the exposure model 
(`expo.mc1`) estimates.

```{r plot-mc1}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$expo.mc1)
```

Plot the empirical cumulative distribution function (ecdf) of the risk model 
(`risk`) estimates.

```{r plot-risk}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk)
```

So if we did this right, could the plot above can be correctly labeled as follows?



```{r plot-riskwlab}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$risk, main = "", ylab = "Proportion of Estimates", xlab = "Daily Probability of Infection")
title("Daily Infec. Risk with Cryptosporidium by RCRW
in the USVI", cex.main = 1, font.main= 1, line = 7)
```


```{r plot-yearriskwlab}
# Generate an "ecdf" plot. This actually calls plot.mcnode().
plot(expo.ev1$yearrisk, main = "", ylab = "Proportion of Estimates", xlab = "Yearly Probability of Infection")
title("Yearly Infec. Risk with Cryptosporidium by RCRW
in the USVI", cex.main = 1, font.main= 1, line = 7)
```


Manually remake the graph, this time using yearrisk, which should just be another mcnode object with all values between 0 and 1:

The first one plots yearrisk without a log x scale:
```{r plot-withyearrisk}
library(reshape)
library(ggplot2)
Risk.mat.1 <- expo.ev1$yearrisk[,,]
probs1 <- c(0.025, 0.25, 0.50, 0.75, 0.975)
quant1 <- as.data.frame(t(apply(Risk.mat.1, 1, quantile, probs = probs1)))
ecdfs1 <- sapply(names(quant1), function(q) ecdf(quant1[[q]]))
plot(ecdfs1[['50%']], main = '')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
lines <- mapply(function(e, g) lines(e, col = g), ecdfs1, grays)
```


Plotting yearrisk ECDF plot with ggplot2:

```{r plot-withyearriskgpplot}
quant.melt1 <- suppressMessages(melt(quant1))
names(quant.melt1) <- c('q', 'x')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
ggplot(quant.melt1, aes(x = x)) + theme_bw() + theme(legend.position = 'none') + geom_hline(yintercept = c(0, 1), linetype = "dashed", color = 'gray') + stat_ecdf(aes(group = q, color = q)) + xlab('x') + ylab('Fn(x)') + scale_colour_manual(values = grays)
```

Plot Version 2 with ggplot2, and a log-scale x axis:

```{r plot-withyearrisklogscale}
quant.melt1 <- suppressMessages(melt(quant1))
names(quant.melt1) <- c('q', 'x')
grays <- c('gray75', 'gray35', 'black', 'gray35', 'gray75')
ggplot(quant.melt1, aes(x = x)) + theme_bw() + theme(legend.position = 'none') + geom_hline(yintercept = c(0, 1), linetype = "dashed", color = 'gray') + stat_ecdf(aes(group = q, color = q)) + xlab('x') + ylab('Fn(x)') + scale_colour_manual(values = grays) + scale_x_log10()
```

